Distributed Algorithms for Sparse Matrix-Vector Multiplication

### 2015-04-28 presentation.Rmd
Some small tweaks

But, for matrices that are **sparse** (contain a lot of zeroes), this is not ideal:
- Waste of memory
- Waste of computation time - Lots of multiplications and additions of zero

@@

@@

## R Simulation
```{r, eval=FALSE}
wearables <- read.csv("data/Example_WearableComputing_weight_lifting_exercises_biceps_curl_variations.csv")
X <- apply(wearables[,-1], 2, as.numeric)
y <- as.numeric(wearables[,1])

@@

t(X) %*% y
t(biggerX) %*% biggery
```

@@

## Algorithms Tested
1. **Block** - no attempt at any load balancing is done.
2. **Cyclic** - Each row i is allocated to processor i mod P, where P is the number of processors.
3. **Aliaga** - An iterative load balancing algorithm that generates swaps of matrix rows among processors to gradually smooth the maxima and minima of load.
4. **Gala** - The algorithm presented in this paper.

@@

## Distributed Memory Parallel Computers (DMPC)
<img src="images/dmpc.png" width="600px">

@@

## The Recursive Algorithm
<img src="images/bil.png" width="500px">

## Step D
<img src="images/stepd.png" width="500px">

## Step C
<img src="images/stepc.png" width="500px">

>>>>

But, for matrices that are **sparse** (contain a lot of zeroes), this is not ideal:

- Waste of memory
- Waste of computation time 
- Lots of multiplications and additions of zero

@@

print(xtable(matrix(rbinom(15, 1, .2), ncol = 3)), type = "html", comment = FALSE)
```

@@

## R Simulation
```{r, warning=FALSE}
wearables <- read.csv("data/Example_WearableComputing_weight_lifting_exercises_biceps_curl_variations.csv")
X <- apply(wearables[,-1], 2, as.numeric)
y <- as.numeric(wearables[,1])

@@

system.time(t(X) %*% y)
system.time(t(biggerX) %*% biggery)
```

@@

## Algorithms Tested
1. **Block** - no attempt at any load balancing is done.
2. **Cyclic** - Each row i is allocated to processor i mod P, where P is the number of processors.
3. **Aliaga** - An iterative load balancing algorithm that generates swaps of matrix rows among processors to gradually smooth the maxima and minima of load.
4. **GALA** - The algorithm presented in this paper.

@@

## Distributed Memory Parallel Computers (DMPC)
<img src="images/dmpc.png" width="400px">

@@

## The Recursive Algorithm
<img src="images/bil.png" width="400px">

## Step D
<img src="images/stepd.png" width="400px">

## Step C
<img src="images/stepc.png" width="400px">
